
\documentclass[11pt,a4paper,computermodern]{article}


\usepackage[
%	includeheadfoot,
	width=172mm,
	top=18mm,
	bottom=18mm,
	bindingoffset=4mm
	]{geometry}



%%% Typeface packages
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{fontawesome}

\usepackage{multirow}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[flushleft]{threeparttable}
\usepackage{enumitem}

%%% Graphics packages
\usepackage{graphicx}
\usepackage{subcaption}


%%

\usepackage{hyperref}
\hypersetup{
	colorlinks=false,
	hidelinks=true,
}


%%
\title{Spotify Data Governance Framework Implementation}
\date{}


\begin{document}

\maketitle

\vspace{-10mm}

This document outlines how Spotify should operationalize its Data Governance Framework, including the organizational model, tools, pilot plan, and execution steps.

\section*{Organizational Model}

Given the size of the company and the amount of data to manage, the recommended approach is that of the center of excellence (CoE) model. In this model, a central data governance team, lead by the CDO, defines the standards and practices for data management. It also provides training and support to the data teams embedded in each department (e.g. marketing, product, engineering). A cross-functional governance committee of data engineers, legal councils, data protection officer and business leaders oversights data management and compliance.

In the context of this organizational model, both full-code and low/no-code solutions will be used.
\begin{itemize}
	\item \textbf{Full-code} technologies are mandatory among the CoE team. They must also be available among the embedded teams to allow for interfacing data manipulation with the CoE team.
	\item \textbf{Low/No-code} technologies are key to the embedded teams in the different department. They provide high data availability for a large volume of non-tech people such as business analysts.
\end{itemize}


\section*{Technologies and tools}

\begin{table}[h]
	\centering
	\begin{threeparttable}
		\caption{Recommended technology tools to support Spotify data governance framework.}
		\label{table:tools}
		\begin{tabularx}{0.99\textwidth}{c >{\centering\arraybackslash}X >{\centering\arraybackslash}X}
			\toprule
			Category & Purpose & Examples \\
			\midrule
			Data cataloging & Central metadata management, data discovery & \href{https://www.alation.com/}{Alation}, \href{https://atlas.apache.org/}{Apache Atlas}, \href{https://atlan.com/}{Atlan}, \href{https://www.collibra.com/}{Collibra} \\
			Data quality & Automate data profiling, validation, alerts & \href{https://www.ataccama.com/}{Ataccama}, \href{https://www.informatica.com/products/data-quality.html}{Informatica Data Quality}, \href{https://www.soda.io/}{Soda}, \href{https://www.talend.com/products/data-integrity-governance/}{Talend} \\
			Compliance monitoring & Manage user consent, automate DSARs & \href{https://bigid.com/}{BigID}, \href{https://www.onetrust.com/}{OneTrust}, \href{https://trustarc.com/}{TrustArc}, \href{https://verasafe.com/}{VeraSafe} \\
			Data security & Protection of sensitive user and payment data & \href{https://www.dataguard.com/}{DataGuard}, \href{https://www.splunk.com/}{Splunk}, \href{https://cpl.thalesgroup.com/encryption/vormetric-data-security-platform}{Vormetric} \\
			Lineage and integration & Track data flow across systems & \href{https://atlas.apache.org/}{Apache Atlas}, \href{https://www.informatica.com/products/data-quality.html}{Informatica} \\
			\bottomrule
		\end{tabularx}
	\end{threeparttable}
\end{table}

It is of course recommended to avoid redundancy by selecting a single tool for a given purpose.


\section*{Pilot implementation}

\subsection*{Overview of the Pilot Implementation}

\textbf{Objective of the Pilot:}\\
To validate the effectiveness of Spotify’s new Data Governance Framework by applying it to the User Engagement Data domain. The rationale behind this choice is that user data is central to key components of the governance plan: compliance, quality and security. This dataset includes user interactions such as play history, search behavior, and playlist creation. The pilot aims to enhance data quality, ensure GDPR/CCPA compliance, and improve internal accessibility and reliability of engagement insights.

\textbf{Scope of the Pilot:}\\
The pilot will focus on the User Engagement Data used by the Product, Marketing, and Analytics teams to assess content performance and drive personalization strategies.


\subsection*{Key Goals}

\begin{itemize}[itemsep=5pt, parsep=0pt]
	\item \textbf{Improve Data Quality} (accuracy, completeness, consistency) in engagement records
	\item \textbf{Ensure Compliance} with GDPR/CCPA for user behavior data
	\item \textbf{Streamline Access} to clean, governed engagement data for authorized staff
	\item \textbf{Mitigate Risks} associated with personal data misuse or poor data quality
\end{itemize}


\subsection*{Pilot Team and Roles}

\begin{table}[ht]
	\centering
	\begin{threeparttable}
		%\caption{}
		\label{table:roles}
		\begin{tabularx}{0.99\textwidth}{c >{\centering\arraybackslash}X >{\centering\arraybackslash}X}
			\toprule
			Team member & Role & Responsibilities \\
			\midrule
			- & Pilot Project Manager & Oversees pilot execution, coordination, and reporting \\
			- & Data Steward & Manages quality and documentation of engagement data \\
			- & Data Protection Officer (DPO) & Conducts privacy audits and ensures GDPR/CCPA compliance \\
			- & Data Analyst Engineer & Implements quality monitoring and access control tools \\
			- & Product Department Lead & Aligns implementation with product strategy and insights \\
			\bottomrule
		\end{tabularx}
	\end{threeparttable}
\end{table}


\subsection*{Timeline and Milestones}

\begin{table}[ht]
	\centering
	\begin{threeparttable}
		%\caption{}
		\label{table:milestones}
		\begin{tabularx}{0.99\textwidth}{c >{\centering\arraybackslash}X >{\centering\arraybackslash}X}
			\toprule
			Milestone & Target date & Responsible \\
			\midrule
			Kick-off meeting & asap & Project Manager \\
			Data Assessment and Cleansing & +1 month & Data Steward \\
			GDPR/CCPA Compliance Audit & +1 month & DPO \\
			Technical Setup and Integration & +2 month & IT Engineer \\
			Mid-Project Review & +2 months & Project Manager \\
			Final Review and Closure & +4 months & Project Manager \\
			\bottomrule
		\end{tabularx}
	\end{threeparttable}
\end{table}


\subsection*{Key Deliverables}

\begin{itemize}[itemsep=5pt, parsep=0pt]
	\item \textbf{Data Quality} Report (before/after metrics for missing/duplicate/inconsistent data)
	\item \textbf{Compliance Assessment} of engagement data processing practices
	\item \textbf{Integration Plan} for user engagement data across product and analytics teams
	\item \textbf{Risk Assessment} Report for data security and exposure
	\item \textbf{Stakeholder Feedback Summary} with insights from product managers and analysts
\end{itemize}


\subsection*{Key Performance Indicators (KPIs)}

\begin{itemize}[itemsep=5pt, parsep=0pt]
	\item 100\% GDPR/CCPA consent tracking on the dataset
	\item 20\% faster access to engagement insights via data catalog
	\item 0 security incidents involving personal data during the pilot
	\item 90\% service-level agreement compliance on data subject requests
%	\item 80\% data literacy training adoption in pilot teams
	\item 10\% reduction in missing or inaccurate user interaction logs
	\item less than 3\% data quality error rate in governed domain
\end{itemize}


\subsection*{Risk Management}

\begin{table}[h]
	\centering
	\begin{threeparttable}
		%\caption{}
		\label{table:risks}
		\begin{tabularx}{0.99\textwidth}{>{\centering\arraybackslash}X >{\centering\arraybackslash}X >{\centering\arraybackslash}X >{\centering\arraybackslash}X}
			\toprule
			Risk & Likelihood & Impact & Mitigation Strategy \\
			\midrule
			Non-compliance with regulations & Medium & High & Weekly audits and DPO signoff \\
			Team resistance to new processes & High & Medium & Stakeholder workshops and clear benefit framing \\
			Metadata gaps in engagement data & Medium & High & Pre-pilot documentation sprint \\
			Technical integration complexity & Medium & High & IT involvement in early scoping \\
			\bottomrule
		\end{tabularx}
	\end{threeparttable}
\end{table}


\subsection*{Training and Change Management}

\begin{itemize}[itemsep=5pt, parsep=0pt]
	\item \textbf{Training Sessions:} Conduct onboarding for Product and Analytics teams
	\item \textbf{Resources:} Share data governance guides, FAQs, and a support channel
	\item \textbf{Feedback Loop:} Weekly check-ins and a post-pilot survey to gather input from users
\end{itemize}


\subsection*{Evaluation and Lessons Learned}

\begin{itemize}[itemsep=5pt, parsep=0pt]
	\item \textbf{Evaluation Metrics:} Compare pre/post KPI results and stakeholder feedback
	\item \textbf{Lessons Learned:} Identify areas of resistance, gaps in metadata, or unanticipated integration issues
	\item \textbf{Feedback Summary:} Compile suggestions to refine the framework before full rollout
\end{itemize}


\subsection*{Next Steps}

\begin{itemize}[itemsep=5pt, parsep=0pt]
	\item \textbf{Scaling Plan:} Expand framework to cover Marketing Data and Content Metadata
	\item \textbf{Adjustments:} Update metadata standards and training materials based on pilot outcomes
	\item \textbf{Full Rollout Proposal:} Present refined strategy and timeline to the Data Governance Committee
\end{itemize}

%\subsection{nice}
%
%Spotify will begin with a pilot implementation in the user data domain. The rationale behind this choice is that user data is central to key components of the governance plan: compliance, quality and security. The steps of this pilot implementation are given below. They should be completed in 3-6 months.
%\begin{itemize}[itemsep=5pt, parsep=0pt]
%	\item[\textbf{1.}] Select Stakeholders: CDO, Data Stewards (Marketing, Product), DPO, Legal.
%	\item[\textbf{2.}] Audit Existing Data: Identify sources, access logs, quality gaps.
%	\item[\textbf{3.}] Apply Governance Rules: Define metadata, access controls, validation checks.
%	\item[\textbf{4.}] Configure Tools: Deploy data catalog and privacy software for pilot scope.
%	\item[\textbf{5.}] Train Teams: Deliver sessions on new policies, roles, and tools.
%	\item[\textbf{6.}] Monitor Metrics: \% of high-quality user records, of DSARs handled on time, user trust/engagement metrics
%	\item[\textbf{7.}] Review \& Iterate: Adjust processes before scaling to other domains.
%\end{itemize}
%
%The pilot implementation will have the following target performance indicators:
%\begin{itemize}[itemsep=5pt, parsep=0pt]
%	\item 90\% service-level agreement compliance on data subject requests
%	\item 80\% data literacy training adoption in pilot teams
%	\item less than 3\% data quality error rate in governed domains
%\end{itemize}


\end{document}